% Copyright 2022 Néstor Nápoles López

One of the most common ways to analyze a piece of tonal
music is through \gls{rna}. This requires the inspection of
several attributes related to chords and keys. For example,
the inversion of a chord, its root, and whether it tonicizes
a momentary musical key. Clearly, if a \gls{rna} is computed
automatically, these tonal attributes become automatically
available too. However, the Roman numeral notation itself
also describes complex---sometimes exotic---chords,
fluctuations of musical key, and other tonal situations in a
very compact syntax. This is maybe the reason why a number
of musicians have preferred it throughout the years. It is a
notation that is compact enough to be annotated within a
regular musical score, yet encodes explanations to advanced
musical concepts that would require much more words and
effort to be described otherwise.

In order to illustrate this more visually, consider the
musical example in \reffig{op28no20_intro_analysis}. In this
example, which features the entirety of Chopin's Prelude Op.
28 No. 20, two successions of chords can be seen in measures
2 and 12. Something interesting about these two successions
is that they start with the same two chords. Namely, an
$\pitchAb$ major triad, and a $\pitchDb$ major triad.
However, the musical key in both instances can be
interpreted differently. 

In the first occurrence (i.e., m.~2), these chords can be
analyzed in the context of the $\keyAb$ major key. The first
chord ($\pitchAb$ major) is a tonic chord, the second chord
($\pitchDb$ major) is a subdominant chord, the chord that
follows is a dominant seventh chord ($\pitchEb\rnseven$),
and the last chord is a tonic triad again ($\pitchAb$
major). In \gls{rna}, we express this as in
\refeq{succession1}.

\begin{equation}
    \label{eq:succession1}
    \rnwkey{\keyAb}{\rnI} \, \rnIV \, \rnV\rnseven \, \rnI
\end{equation}

In the second occurrence (i.e., m.~12), the chords can be
analyzed in the context of the $\keyc$ minor key, which is
the key of the piece. Here, the first chord ($\pitchAb$
major) is a submediant chord, the second chord ($\pitchDb$
major) is a flattened second degree, commonly known as a
\gls{neapolitan} chord, the third chord is a dominant
seventh chord ($\pitchG\rnseven$), and the last chord is the
tonic of the piece $\pitchC$ minor.

\begin{equation}
    \label{eq:succession2}
    \rnwkey{\keyc}{\rnVI} \, \musFlat{}\rn{II} \, \rnV\rnseven \, \rnI
\end{equation}

\phdfigurefullpage[Musical analysis of two successions of
chords within Chopin's Op.~28
No.~20.]{op28no20_intro_analysis}