% Copyright 2022 Néstor Nápoles López

% This is \refsubsubsec{thedcmlstandard}, which introduces
% the the dcml standard.

This standard for harmonic annotations was developed at the
\gls{dcml} at the \'Ecole Polytechnique F\'ed\'erale de
Lausanne.

This standard was first used in the annotation of the
\gls{abc}
\parencite{neuwirth2018annotated}. Since then, it has been
revised and documented in a dedicated
repository\footnotelink{https://github.com/DCMLab/standards}
and a reference
manual.\footnotelink{https://dcmlab.github.io/standards/build/html/reference/reference.html}

The introduction section of a tutorial for this annotation
standard summarizes the philosophy behind the standard:

\begin{italicquote}
    The goal of this tutorial is to provide a systematic and
    condensed way of conveying the annotation philosophy and
    principles (aka “the guidelines”) behind the annotation
    standard. What has been condensed are more than two
    years of discussions between the music theory experts
    involved in the standard's creation and application,
    i.e. between annotators, reviewers, and users.
    Therefore, this tutorial is not about telling anybody
    “how harmonic analysis really works” or “how everyone
    should be using Roman numerals”. Instead, it introduces
    a set of guidelines for analysts who want to use the
    DCML harmony annotation standard to encode a set of
    musical features in a consistent and machine-readable
    manner so that others can re-use and rely on the encoded
    information.
\end{italicquote}

In addition to the repository and reference manual, the
community behind this standard has also made available the
corresponding annotation tutorial mentioned
above,\footnotelink{https://dcmlab.github.io/standards/build/html/tutorial/index.html},
an annotation workflow that discusses version control
practices,\footnotelink{https://dcmlab.github.io/standards/build/html/git/git.html},
a set of instructions for corpus creation (described by
\gls{dcml} as the corpus-creation
pipelin),\footnotelink{https://dcmlab.github.io/standards/build/html/pipeline/pipeline.html},
and a page of questions and
examples,\footnotelink{https://dcmlab.github.io/standards/build/html/reference/examples.html},
among other resources.


From the perspective of documentation and standardization,
this is perhaps the most thoroughly described standard for
\gls{rna} in existence today.

The revised version of the standard has been used in a
subsequent iteration of the \gls{abc} dataset (version 2) as
well as the \gls{mps} dataset
\parencite{hentschel2021annotated}. Other collections exist
in a private archive and are expected to be publicly
available one day.
