\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\label{chap:acknowledgements}

I will first and foremost thank my PhD advisor, Ichiro
Fujinaga. For the most part of my PhD, Ichiro scheduled
weekly meetings with me that were of the utmost help.
Together, we walked through machine-learning prototypes,
presentations, code, music theories, and papers. Ichiro and
Masako also made me feel supported since I set foot in
Montr\'eal,\footnote{To be precise, since I set foot in
Canada for the first time in my life, at the Trudeau airport
in Montreal.} making it a pleasant experience for me and my
partner to live and work in the city all these years. Thank
you.

Secondly, I would like to thank my family, who have
supported me all along: my spouse and best friend, Kinga; my
parents, Martha and Eduardo; my siblings, Grecia, Eric, and
sister-in-law Fabiola; and my parents-in-law, Gra\.zyna and
Miros\l{}aw.

After my advisor, my co-authors contributed me with many
ideas and feedback related to the research presented here.
Thanks to all of you: Gabriel Vigliensoni, Claire Arthur,
Laurent Feisthauer, Florence Lev\'e, Mark Gotham, and the
\gls{cts} group.\footnote{Jacob deGroot-Maggetti, Laurent
Feisthauer, Ichiro Fujinaga, Sam Howes, Yaolong Ju, Suzuka
Kokubu, Sylvain Margot, Tim de Reuse, and Finn Upham.}

Thanks to all the members of the \gls{ddmal} and
\gls{cirmmt} communities, but particularly, thanks to
Gabriel Vigliensoni, Martha Thomae, Matan Gover, Yulia
Draginda, Yuval Adler, Alex Daigle, Emily Hopkins, Yaolong
Ju, Timothy de Reuse, Sylvain Margot, Jacob Sanz-Robinson,
Erica Huynh, and K\'e Zhang.

Thanks to McGill professors Julie Cumming, Peter Schubert,
William Caplin, Jon Wild, Philippe Depalle, Marcelo
Wanderley, Gary Scavone, and Stephen McAdams from whom I
learned different aspects of music and technology that
contributed explicitly or implicitly to the work presented
in this dissertation.

Thanks to Joe Plazak, Edith Gosselin Picard, Rob Gonsalves,
and everyone in the Avid, Sibelius, and RAD Lab families,
who have welcomed me as one of their own, and kindly
mentored me in the art of applied research and
commercial-grade music software development.

Thanks to other members of the global music information
retrieval and computational musicology communities,
particularly, Craig Sapp, Helena Cuesta, Fabian Moss, Emilia
Parada-Cabaleiro, Michael Scott Cuthbert, Jinny Park, Jacob
Tyler Walls, Reinier de Valk, Rebecca Salganik, Johannes
Hentschel, and others with whom I shared code, papers, and
friendship.

The research in this dissertation benefitted from the
generous support of different institutions. The \gls{frqsc},
from which I obtained a doctoral scholarship from 2019--2022
(dossier number 271479); \gls{cirmmt}, which awarded me with
travel funding and a student project award; the \gls{simssa}
project, which provided a generous stipend during the first
two years of my PhD (2017--2018) and a role as a casual
research assistant for the remainder of it; the Schulich
School of Music, which provided additional travel and
conference registration awards; the Digital Research
Alliance of Canada, which facilitated computing resources to
run experiments; and Mitacs, which funded part of an
internship project at Avid Technology. Thank you to all of
these wonderful organizations and their commitment to music
research.

Special thanks to Laurent Feisthauer for translating the
English abstract of this dissertation into French, and to
Genevi\`eve Gates-Panneton for proofreading the final
version of it.

Thanks to people who had an early influence in me and set me
in my current path: Luis Alberto Casillas Santill\'an,
Abelardo Guti\'errez, Juan Jos\'e L\'opez Sandoval, Nancy
Arana Daniel, Sergio Bola\~nos, Leo Hendrik Reyes, and Juan
Rodrigo P\'erez C\'ardenas. Lastly, my friends, for always
being there: Jhonnatan Razo, Dulce Alcal\'a, and El\'i
Lezama. ¡Gracias, amigos!
