\chapter*{Contributions of Authors}
\addcontentsline{toc}{chapter}{Contributions of Authors}
\label{chap:contributionsofauthors}

\paragraph{Chapter 1}
All contributions in this chapter are by \myself{}.

\paragraph{Chapter 2}
All contributions in this chapter are by \myself{}. However,
N\'estor benefitted from a fruitful discussion in
SMTDiscuss,\footnotelink{https://discuss.societymusictheory.org/discussion/553/history-of-roman-numeral-analysis}
which lead to new references and ideas. Particularly, thanks
to Nicolas Mee\`us for participating in this conversation.

\paragraph{Chapter 3} 
\refsubsec{keyestimation} takes several ideas and findings
from a published conference paper
\parencite{napoleslopez2020local}. Particularly, the reviews
of \gls{gke} and \gls{lke} models was conducted by \myself{}
and Laurent Feisthauer, in equal contribution. The review of
\gls{gke} had also a minor contribution by Florence Lev\'e.

\paragraph{Chapter 4} 
During the course of this research, Mark Gotham and
\myself{}, independently (and in collaboration, during the
writing of a co-authored conference paper,
\textcite{napoleslopez2021augmentednet}) contributed
conversions, revisions, and corrections to the listed Roman
numeral analysis datasets. Workflows for data ``hygiene''
and correcting mistranscriptions were also independently
developed by both. However, the data-curation workflows
described in this dissertation are the ones by \myself{}. 

\paragraph{Chapter 5}
This chapter is based on a published conference paper
\parencite{napoleslopez2021augmentednet}. The
\gls{augmentednet} neural network architecture is a
contribution of \myself{}. The data-augmentation technique
based on generating synthetic training examples is a
contribution of \myself{}. Nevertheless, co-author Mark
Gotham first proposed the idea of artificial texturization
of the synthesized examples. The patterns chosen for
texturization and the software implementation are a
contribution of N\'estor N\'apoles L\'opez.

\paragraph{Chapter 6}
This chapter is based on the same published conference paper
as Chapter 5 \parencite{napoleslopez2021augmentednet}. All
the experiments reported in this chapter were conducted by
\myself{}. Likewise for all tables and comparisons.

\paragraph{Chapter 7}
All contributions in this chapter are by \myself{}.
