Roman numeral analysis is a common methodology in
contemporary music theory textbooks dealing with tonal
music. The name derives from the use of Roman numerals to
denote scale degrees, which, in turn, indicate diatonic
chords in a certain key context.

\guide{Why is Roman numeral analysis important compared to, say, chord labels?}
Roman numeral labels provide more information than chord
labels. Generally, a chord label indicates the root of a
chord and its quality. A Roman numeral label provides this
as well, but also information about keys, inversions, and
spellings. For example, Roman numerals are always relative
to a key, so the key must be indicated at all times to
disambiguate the meaning of the Roman numeral. This is a
helpful property for passages of music that feature quick
modulations or changes of key. In the modern syntax, it is
also customary to indicate \emph{applied chords} or
\emph{secondary dominants}. Once the concept of
\emph{tonicization} emerged, in the twentieth century, this
quickly became an important idea of tonal music,
particularly relevant to analyze the chromatic music of the
nineteenth century. Roman numerals facilitate the indication
of modulations as well as tonicizations. With these, a
musical analysis of a particular person can be more
informative, for example, indicating how the analyst
envisions the key context at a particular moment of the
piece. Roman numerals indicate inversions for triads and
seventh chords. Presumably, the Arabic numeral notations
employed to denote inversions in Roman numerals evolved from
figured bass. However, there is generally a small vocabulary
of numerical annotations appended to a Roman numeral, which
indicate the inversion of the chord. There are other subtle
but important differences that come with Roman numeral
analysis compared to chord labels. Think of a diminished
seventh chord, say, F$\sharp$ diminished seventh. The most
usual context of that chord is to anticipate a resolution to
a G minor triad. However, if the chord is in first
inversion, then it is enharmonically equivalent to an A
diminished seventh chord. It is tempting to write the chord
as A diminished, rather than F$\sharp$ with A in the bass,
however, these chords have two very different meanings
contextually. This is evident in the presence of pitch
spelling. The G$\flat$--G$\natural$ results in a strange
voice leading, whereas the F$\sharp$--G is a customary
leading-tone to tonic minor second. These subtle
interactions of spelling are generally honored by the Roman
numeral analysis notation. There are multiple examples that
come to mind, such as Neapolitans, Augmented Sixth Chords,
and \emph{Cadential six-four} chords, where the syntax of
Roman numerals is helpful to provide a more thorough
description of the annotated chord. These differences might
seem negligible to some, but I would like to make a few
arguments: (1) in tonal music theory textbooks, it is more
common to see Roman numeral analysis notation than chord
labels, which might suggest that the more granular notation
is preferred (2) in machine learning contexts, as is the
case for this dissertation, more information in the
annotations can be really helpful, without this additional
information, we could not attempt to learn a key-estimation
model simultaneously using the same annotated data, (3) the
syntax of Roman numeral analysis is highly compressed, it is
generally not more time-consuming to write Roman numerals
than it is to write chord labels, the time-consuming aspect
of Roman numerals is the train of thought required to commit
to a particular interpretation of the analysis (i.e., key
context, chord, inversion, function of the chord), and that
is precisely the thorough thought pattern that we want to
capture in the data.
