% Copyright 2022 Néstor Nápoles López

Throughout the dissertation, various contributions towards
\glspl{rna} have been made in different dimensions of the
problem. \refchap{introductiontoromannumeralanalysis}
presented a historical analysis of the \gls{rna} syntax.
Although several accounts of the history of harmony
exist,\footnote{See, for example,
\textcite{wason1985viennese}, \textcite{grave1988praise},
\textcite{christensen2002tonality},
\textcite{laitz2010graduate}, or
\textcite{sansallovich2013quintas}} most of these do not
focus on the syntax of the annotation. The review presented
in this dissertation focuses solely in the use of symbols,
conventions, and specificities of the \gls{rna} notation
that are relevant for its digitization and standardization.

\refchap{dataacquisitionandpreparation} documents the
aggregation process of the, to the best of my knowledge,
largest publicly available dataset of \glspl{rna} labels.
The process to overcome the alignment between annotations
and scores are discussed through the use of quantitive
metrics and other resources that facilitate the aggregation
of the data. Furthermore, in \refsubsec{preprocesseddata} of
the present chapter, the preprocessed data is made publicly
available for other researchers who may be interested in
using it. The distributed dataset includes the explicit
splits into training, validation, and test portions used for
the experiments conducted in this dissertation. Lastly, in
the context of data, a data augmentation technique was
presented in \refsubsec{synthesisofartificialexamples}.
Although less impactful than transposition, it presents the
advantage that it can be used in combination, increasing the
performance obtained by the network on the same dataset when
using transposition and synthesis of training examples.

\refchap{modeldesign} describes a novel neural network
architecture for \glspl{rna}. The architecture is based on a
previous version presented in
\textcite{napoleslopez2021augmentednet}, however, the
structure of the multitask learning configuration presented
here shows several improvements. Unlike other models, a set
of tasks that have a more balanced distribution of target
classes is proposed for different components. For example,
rather than predicting chord qualities, which are heavily
skewed towards major and dominant seventh examples in the
dataset, chords are modeled by predicting four note
classifiers in the ``SATB'' form of a closed position chord.
This is an equivalent task to predict chords with a more
balanced distribution of classes. Other contributions of the
model include the use of a \gls{pcset121} task, which
summarizes the chord vocabulary of the system, a
\gls{harmonicrhythm7} task based on 7 classes of outputs,
and a vocabulary of 31 Roman numeral numerators \gls{rn31}.
The input representations sent to the neural network also
present contributions. For example, the \gls{duration14}
encoding of note and measure onsets was shown to improve the
segmentation of the chords in the ablation studies shown in
\refsec{bestperformingmodelconfiguration}. The alternative
representation of pitch spelling described in
\refsubsubsec{19two-hot-encoding} was also shown to be an
adequate subtitute for the more common representation in
\refsubsubsec{35one-hotencoding}, without losing
performance.

\refchap{experimentalevaluation} introduced contributions
regarding the evaluation of a \glspl{rna} system. For
example, the performance on rare chords was discussed among
recent approaches in \refsubsec{romannumeralaccuracy}. This
form of evaluation revealed that some of the chords have a
tiny representation in the dataset, however, represent the
chords and tonal situations that most often benefit from the
\gls{rna} notation. In this respect, one of the
contributions is to propose a system that can improve on
such difficult cases, and to signal to future researchers
that methods for evaluating these rare occurrences of chords
are needed in the future.

Lastly, the source code of the model, experiment logs, and
general documentation of the software is offered as a
contribution to future researchers in the field. These
aspects are explained next.
