% Copyright 2021 Néstor Nápoles López

% This is \refsubsubsec{local-keyestimationmodels}, which
% introduces the local-key estimation models.


Contrary to \gls{gke} approaches, \gls{lke} models have a
relatively recent history.

% Audio
\textcite{purwins2000new} introduced a method for tracking
changes of key in audio signals using cq-profiles, which
were calculated with the constant-Q filter bank. Their goal
was to track the tone center and its variation during the
piece. Their dataset included annotations for both
modulations and tonicizations but considered that the ground
truth was the one indicated by the tonicizations.

% symbolic
\textcite{chew2002spiral} measured the distance from a
sequence of pitches to a key using the \emph{Spiral Array}.
The succession of keys was then modeled as a sequence of
\emph{boundaries} dividing the score in different key areas.

% Audio
\textcite{chai2005detection} designed a model based on a
Hidden Markov Model (\gls{hmm}) to detect changes of key.
They described the term \emph{modulation} as ``the change of
key at some point''. Their model detected first the tonal
center and then the mode of the key.

% Audio
\textcite{catteau2007probabilistic} introduced a model for
scale and chord recognition, assuming that a correspondence
between a major scale and a major key, and between a
harmonic minor scale and a minor key. Their model was based
on the key profiles by \textcite{temperley1999whats} and
Lerdahl's \emph{Tonal Pitch Space}
\parencite{lerdahl2005tonal}.

\textcite{izmirli2007localized} proposed a model to find
local keys in audio signals. As a pre-processing step, the
model adapted the tuning of the signal before computing the
spectral analysis and the chroma features. In a later stage,
non-negative matrix factorization was used to segment
contiguous chroma vectors, identifying potential local keys.
The model identified segments that were candidates for
unique local keys in relation to the neighboring key
centers. The model was evaluated in three datasets: 17 pop
songs with at least one modulation, excerpts from the
beginning of 152 classical music pieces from the Naxos
website, and the examples from the
\textcite{kostka1984tonal} harmony textbook. In the paper,
the results for three different evaluation methods were
provided for each of the datasets. Izmirli also attempted to
disambiguate modulations and tonicizations in the following
manner:

\begin{italicquotes}
    Secondary functions and tonicizations are heard as short
    deviations from the well-grounded key in which they
    appear---although the boundary between modulation and
    tonicization is not clear cut. A modulation
    unambiguously instigates a shift in the key center.
\end{italicquotes}

This work is also, to the best of my knowledge, the first
one where the term \emph{local key} was mentioned in
\gls{mir} research.

\textcite{papadopoulos2009local} adopted a similar approach
to \textcite{izmirli2007localized} for audio \gls{lke}.
Their model attempted to segment the score based on the
points of modulation. They introduced key dependencies on
the harmonic and metrical structures of \gls{gke} methods,
in order to convert them into \gls{lke} ones. The method
extended previous work
\parencite{papadopoulos2008simultaneous}, adding a stage of
local key estimation using an \gls{hmm}. The observations of
the model were derived from different key profiles:
\textcite{krumhansl1982tracing},
\textcite{temperley1999whats}, and a flat diatonic key
profile (a uniform distribution for diatonic pitch-classes
and \emph{zero} elsewhere). The system was evaluated with
five Mozart piano sonatas, where the local keys and chords
were manually annotated. The maximum accuracy achieved was
80.22\%

\textcite{rocher2010concurrent} introduced a model that
provided duples of (chord, key) for each audio frame of an
input excerpt. The model was based on a graph and the
\emph{best-path} estimation method. For evaluating key
distances, they used the key profiles by
\textcite{temperley1999whats}. The authors alluded to the
term modulation when discussing their key predictions.

\textcite{mearns2011automatically} used an \gls{hmm} to
estimate modulations over audio transcriptions of Bach
chorales. The \gls{hmm} was trained with chord progressions.
The emission probability distributions were obtained from
two tables with the probabilities of chords existing in a
given key. These tables were based on the work by Schoenberg
and Krumhansl. Tonicizations were not described in these
charts, therefore, the authors did not deal with
tonicizations.

% TODO: Which Schoenberg and Krumhansl?

\textcite{pauwels2014combining} presented a probabilistic
framework for the simultaneous estimation of chords and keys
in audio. They mentioned the importance of ``integrating
prior musical knowledge'' into a \gls{lke} model, however,
they did not allude to the terms modulation and
tonicization. The same year, \textcite{weis2014chromabased}
proposed an audio scale estimator. They argued that this
estimator could help to determine the local tonality based
on G\'{a}rdonyi's scale analysis method. They did not use
the term tonicization, however, they discussed ``short-time
local modulations'', which resemble tonicizations.

Machine learning approaches, especially using neural
networks, have recently gained popularity in \gls{mir}
research, including \gls{lke} models.
\textcite{chen2018functional, chen2019harmony} and
\textcite{micchi2020not} designed models that estimate local
keys as part of their \glspl{rna}
models. Tonicization information was implied by the roman
numeral analysis annotations.

We introduced a \gls{gke} and \gls{lke} model using an
\gls{hmm} \parencite{napoleslopez2019keyfinding}. The model
was capable of working with symbolic and audio data. In that
work, we did not allude to the terms modulation or
tonicization when describing the nature of local keys,
however, an experiment on Chopin's Op.~28 No.~20 revealed
that different key profiles tended to be more or less
sensitive to smaller fluctuations of key (e.g.,
tonicizations). This inspired future work on the
relationship of local keys, modulations, and tonicizations
\parencite{napoleslopez2020local}.

One of the most recent models for finding changes of key is
the one by \textcite{feisthauer2020estimating}, which was
designed to detect modulations in Mozart piano sonatas. It
used three proximity measures established from pitch
compatibility, tonality anchoring, and tonality proximity.
The model computed the cost of being in a key on a given
beat, and estimated the succession of keys using dynamic
programming techniques.
