% Copyright 2022 Néstor Nápoles López

Most tonal music analysis neural network models collapse all
enharmonic spellings of the same note, as shown in
\refeq{enharmonic_equivalence}.

\begin{equation}
    \label{eq:enharmonic_equivalence}
    \pitchFs = \pitchGb \quad \text{enharmonic equivalence}
\end{equation}

\begin{equation}
    \label{eq:enharmonic_nonequivalence}
    \pitchFs \neq \pitchGb \quad \text{enharmonic nonequivalence}
\end{equation}

In \gls{augmentednet}, however, the spelling of a pitch is
encoded and taken into account, as shown in
\refeq{enharmonic_nonequivalence}. Two methods are described
to encode pitch spellings in this way.

%  and encode each of those spelling classes as one
% dimension in a vector. This method has been used in
% \textcite{micchi2020not, micchi2021deep}. An alternative
% method, proposed here, is to encode the pitch class and
% spelling of the pitch class. Both methods are described
% next.

\phdparagraph{35 one-hot encoding}

Perhaps the easiest method to encode a pitch spelling is to
assume that only a number of accidentals are allowed next to
a generic note letter. The use of \gls{sharp} and
\glspl{flat} accidentals is very common. The use of
\glspl{doublesharp} and \glspl{doubleflat} accidentals is
less common. Any accidental sharper than \glspl{doublesharp}
or flatter than \glspl{doubleflat} is rare. Thus, a
reasonable vocabulary for spelled pitch
classes\footnote{Note that the octave is ignored.} is one
with 35 classes, from $\pitchCbb$ to $\pitchBx$, as shown in
\refeq{pitch_spelling_35}. A ``one-hot'' encoding of this
form means that, at each timestep, one out of the 35
possible classes of pitch spelling will be set as active
(``hot'') and the rest will not.

\begin{equation}
    \label{eq:pitch_spelling_35}
    \begin{split}
    \setps{}_{35} = \{ & \pitchCbb, \pitchDbb, \pitchEbb, \pitchFbb, \pitchGbb, \pitchAbb, \pitchBbb, \\
    & \pitchCb, \pitchDb, \pitchEb, \pitchFb, \pitchGb, \pitchAb, \pitchBb, \\
    & \pitchC, \pitchD, \pitchE, \pitchF, \pitchG, \pitchA, \pitchB, \\
    & \pitchCs, \pitchDs, \pitchEs, \pitchFs, \pitchGs, \pitchAs, \pitchBs, \\
    & \pitchCx, \pitchDx, \pitchEx, \pitchFx, \pitchGx, \pitchAx, \pitchBx \}
    \end{split}
\end{equation}

This method was used in \textcite{micchi2020not,
micchi2021deep} to encode the inputs of two \gls{arna}
models. Using this approach, a 35-dimensional pitch-spelling
vector is encoded every timestep, indicating the active
pitch (or pitches, if several) among the 35 available
classes. One of the limitations of this method is that it
cannot encode any notes sharper than a \glspl{doublesharp}
or flatter than a \glspl{doubleflat}. Additionally, the
resulting vector is almost three times larger than an
encoding based on pitch classes, which only requires 12
classes,\footnote{When enharmonic equivalence is assumed in
a \gls{12tet} system, a set of 12 pitch classes spans all
the note classes of the Western chromatic scale.} as shown
in \refeq{pitch_classes}.

\begin{equation}
    \label{eq:pitch_classes}
    \mathcal{C} = \{0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11\}
\end{equation}

\phdparagraph{19 two-hot encoding}


Another encoding method, first proposed in
\textcite{napoleslopez2021augmentednet}, provides an
alternative way to encode pitch spellings. Instead of
encoding \gls{sharp} and \glspl{flat} notes explicitly, it
extends the pitch class vector representation, by
concurrently encoding a generic note letter class, as shown
in \refeq{pitch_spelling_19}.

\begin{equation}
    \label{eq:pitch_spelling_19}
    \begin{split}
        \setletters &= \{ \pitchC, \pitchD, \pitchE, \pitchF, \pitchG, \pitchA, \pitchB \} \\ 
        \setps_{19} &= \mathcal{C}
        \times \setletters
    \end{split}
\end{equation}

% \begin{equation} \label{eq:pitch_spelling_19_magnitude} |
%     \setps_{19} | = 19 \end{equation}

This encoding method results in duples of the form $(c, l)$,
where $c \in \mathcal{C}$ and $l \in \setletters$. For
example, C\musSharp{} is represented by the duple $(1,
\text{C})$, whereas D\musFlat{} is represented by the same
pitch class but a different generic note letter, $(1,
\text{D})$. D$\musNatural{}$ has the same generic note
letter but a different pitch class, $(2, \text{D})$, and so
on. Using this representation, a spelled note, beyond two
flats and two sharps can be encoded with a 19-feature
vector. Notice, however, that this vector is a two-hot
encoding representation. This means that at any given
timestep, each pitch spelling will be encoded as two active
classes: one for the pitch class and one for the generic
note letter. The output representations in the \gls{mtl}
configuration (see \refsec{multitasklearningoutputs}) use
the first encoding method. The input representations
described in \refsubsec{inputrepresentations} employ the
second encoding method. 
