% Copyright 2022 Néstor Nápoles López

In order to perform the evaluation, all the annotations in
the \gls{romantext} files, including the ground truth, are
decomposed into their \gls{pcset} $\elpcset$, key $\elkey$,
and inversion $\elinv$ components. The components $\elpcset$
and $\elkey$ are sent to the \gls{natem} algorithm to
retrieve a numerator $\elnum$ and optional tonicization
$\elden$. This process is illustrated for an example of the
test set in \reftab{natem_translation}. The process
performed on the predictions of each model is identical.

\phdtable[Translation process of the ground-truth
annotations in one of the test files in the \gls{kmt}
dataset. The original annotations (leftmost column) are
parsed from a \gls{romantext} file, and decomposed into a
\gls{pcset} $\elpcset$, key $\elkey$, and inversion $\elinv$
components. Furthermore, a numerator $\elnum$ is retrieved
from \gls{natem} (fifth column). In some cases, \gls{natem}
will also return a tonicization, corresponding to the
rightmost column]{natem_translation}

In the evaluation of the models for accuracy on individual
tasks, the components $\elpcset$, $\elkey$, and $\elinv$ are
compared to the ground truth. In the Roman numeral accuracy,
the $\elnum$ numerator of the ground truth is compared
against the $\elnum$ of the models. This second evaluation
is helpful to find how well the models are able to retrieve
the different tonal contexts of the chords in the ground
truth. These results are shown in the form of confusion
matrices of the target numerator $\elnum$ against the
predicted numerator $\elnum$.


