% Copyright 2022 Néstor Nápoles López

In order to perform the evaluation, all the annotations in
the \gls{romantext} files, including the ground truth, were
decomposed into their \gls{pcset} $\elpcset$, key $\elkey$,
and inversion $\elinv$ components. The components $\elpcset$
and $\elkey$ were processed by \gls{natem} to retrieve a
numerator $\elnum$ and tonicization $\elden$. This process
is illustrated for a ground-truth example of the test set in
\reftab{natem_translation}. The process performed on the
predictions of each model was identical, using their
\gls{romantext} files.

\phdtable[Translation process of the ground-truth
annotations in one of the test files of the \gls{kmt}
dataset. The original annotations (leftmost column) are
parsed from a \gls{romantext} file, and decomposed into a
\gls{pcset} $\elpcset$, key $\elkey$, and inversion $\elinv$
components (middle columns). Then, a numerator $\elnum$ and
tonicization $\elden$ were  retrieved from \gls{natem}
(rightmost columns)]{natem_translation}

Three evaluation experiments were performed. The first one,
measuring the time it took for the models to generate their
chord predictions. Two subsequent experiments evaluated the
performance of the models on their \gls{rna} predictions:

\begin{enumerate}
    \item The models were evaluated in the accuracy of their
    $\elpcset$, $\elkey$ and $\elinv$ components, compared
    to the $\elpcset$, $\elkey$ and $\elinv$ of the
    ground-truth.
    \item The models were also evaluated using the $\elnum$
    retrieved from \gls{natem}, which was compared against
    the $\elnum$ obtained for the ground truth.
\end{enumerate}


% In the evaluation of the models for accuracy on individual
% tasks, the components $\elpcset$, $\elkey$, and $\elinv$
% are compared to the ground truth. In the Roman numeral
% accuracy, the $\elnum$ numerator of the ground truth is
% compared against the $\elnum$ of the models. This second
% evaluation is helpful to find how well the models are able
% to retrieve the different tonal contexts of the chords in
% the ground truth. These results are shown in the form of
% confusion matrices of the target numerator $\elnum$
% against the predicted numerator $\elnum$.


